---
title: "Validation of IBM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(cowplot))
library(ICDMM)
```

# Equilibrium testing

This vignette demonstrates that the new individual-based model (IBM) run through 'malariasimulation' is consistent with the [equilibrium solution](https://github.com/mrc-ide/malariaEquilibrium), legacy IBM, and [compartmental malaria model](https://github.com/mrc-ide/deterministic-malaria-model).

We can load some runs from the legacy model and store the EIRs for comparison.

```{r}
old_model_runs <- readRDS(
  system.file('extdata', 'eir_prev_inc_old_model.RDS', package='msvalidation')
)
old_model_runs$model <- 'legacy'
eir <- old_model_runs$eir
```

Let's start by running the malariasimulation for a range of EIRs:

```{r}
year <- 365
warm_up <- 5 * year
period <- 5
population <- 1e5
eta <- 0.0001163315

fillna <- function(x, value) {
  x[is.na(x)] <- value
  x
}

# get_ms_runs <- function(eir) {
#   params <- malariasimulation::get_parameters(list(
#     human_population = population,
#     clinical_incidence_rendering_min_ages = 0,
#     clinical_incidence_rendering_max_ages = 5 * year,
#     individual_mosquitoes = FALSE,
#     n_heterogeneity_groups = 5,
#     average_age = 1 / eta
#   ))
#   params <- malariasimulation::set_equilibrium(params, eir)
#   malariasimulation::run_simulation(warm_up + period * year, params)
# }
# 
# get_ms_values <- function(run) {
#   output <- tail(run, period * year)
#   c(
#     total_m = mean(output$total_M) / population,
#     clinical_incidence = mean(fillna(output$clin_inc_0_36500, 0)) * year,
#     clinical_incidence_0_5 = mean(fillna(output$clin_inc_0_1825, 0)) * year,
#     prevalence_2_10 = mean(output$pv_730_3650),
#     eir = mean(output$EIR) * year
#   )
# }
# 
# ms_runs <- lapply(eir, get_ms_runs)
# ms_values <- data.frame(
#   t(
#     vapply(
#       ms_runs,
#       get_ms_values,
#       numeric(5)
#     )
#   )
# )
# ms_values$model <- 'malariasimulation'

# saving malariasimulation runs for future ref
#saveRDS(ms_runs,"./inst/ms_runs.rds")
#saveRDS(ms_values, "./inst/ms_values.rds")
```

And compute some outputs from malariaEquilibrium...

```{r}
eq_params <- malariaEquilibrium::load_parameter_set()
eq_params$eta <- eta
init_age <- c(seq(0,2,by=0.25),seq(2.5,6,by=0.5),seq(7,15,by=1),17.5,20,seq(25,100,by=5))

get_eq <- function(eir) {
  malariaEquilibrium::human_equilibrium(
    EIR = eir,
    ft = 0,
    p = eq_params,
    age = init_age
  )
}

get_equilibrium_values <- function(eq, eir) {
  u5 <- init_age < 5 # equilibrium model output gives a matrix of values for each age
  g2l10 <- (init_age >= 2) & (init_age < 10)
  c(
    total_m = NA,
    clinical_incidence = sum(eq$states[, 'inc']) * year,
    clinical_incidence_0_5 = sum(eq$states[u5, 'inc']) / sum(eq$states[u5, 'prop']) * year,
    prevalence_2_10 = sum(eq$states[g2l10, 'pos_M']) / sum(eq$states[g2l10, 'prop']),
    eir = eir
  )
}

eq_data <- data.frame(
  t(
    vapply(
      old_model_runs$eir,
      function(e) get_equilibrium_values(get_eq(e), e),
      numeric(5)
    )
  )
)
eq_data$model <- 'equilibrium'
```

And compute some outputs from deterministic-malaria-model...

```{r}
ms_runs <- readRDS("../inst/ms_runs.rds")
ms_values <- readRDS("../inst/ms_values.rds")

# creating a function for running the odin model given an assigned EIR value
get_dmm <- function(eir) {
  run_model(model = 'odin_model_ageprev',
                     time = period*year,
                     age = init_age,
                     #age = c(0, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.25, 4.5, 4.75, 5, 7.5, 10, 15, 20, 30, 40, 50, 60, 70, 80),
                     het_brackets = 5,
                     init_EIR = eir,
                     init_ft = 0,
                     eta = eta,
                     agestart = 9, #2
                     ageend = 20) #22=7.5
            
} 

# creating a function to output CI and Prev values for a given EIR value
get_dmm_values <- function(eq,eir) {
  c(
    total_m = NA,
    clinical_incidence = mean(eq$inc) * year, # DMM outputs daily CI
    clinical_incidence_0_5 = mean(eq$inc05) * year, 
    prevalence_2_10 = mean(eq$prev_agegroup), # prev output is for ages
    eir = eir
  )
}

# loop through functions, output a data frame with CI and Prev values for each EIR value
dmm_data <- data.frame(
  t(
    vapply(
      old_model_runs$eir, 
      function(e) get_dmm_values(get_dmm(e), e),
      numeric(5)
    )
  )
)

# labeling model outputs
dmm_data$model <- 'deterministic'

```

We can then compare our malariasimulation outputs to the equilibrium prevalence and incidence curves by EIR:

```{r, fig.height = 5, fig.width = 7}
# consolidate to a single data frame
out <- rbind(old_model_runs, eq_data, ms_values, dmm_data)

# plot
plots <- list(
  ggplot(
    out,
    aes(x = eir, y = prevalence_2_10, group = model, color = model)
  ) + 
    geom_line(col = "grey") + 
    geom_point() + 
    scale_x_log10() +
    theme_bw() +
    labs(title = 'Adult EIR vs Prev'),
  ggplot(
    out,
    aes(x = eir, y = clinical_incidence_0_5, group = model, color = model)
  ) + 
    geom_line(col = "grey") + 
    geom_point() + 
    scale_x_log10() +
    theme_bw() +
    labs(title = 'Adult EIR vs Inc')
)

plot_grid(plotlist = plots)
```

We can then look at the expected state counts for a medium EIR:

```{r, fig.height = 12, fig.width = 7}
agreement_plot <- function(output, eq, title='Agreement') {
  title <- ggdraw() + 
    draw_label(
      title,
      fontface = 'bold',
      x = 0,
      hjust = 0
    ) +
    theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0, 7)
    )
  plot_grid(
    title = title,
    plotlist=lapply(
    c('S', 'D', 'A', 'U'),
    function(state) {
      ggplot(output) +
        geom_line(aes_string(x = 'timestep', y = paste0(state, '_count'))) +
        geom_hline(
          yintercept = sum(eq$states[,state]) * population,
          linetype = 'dashed'
        ) +
        labs(title = state)
    }
  ))
}

# plotting 
agreement_plot(
  ms_runs[[5]],
  get_eq(eir[[5]]),
  title=paste0('EIR = ', round(eir[[5]], 2))
)
```
